Floral choice

### 2020-12-10 analyses/floral_choice.Rmd
anova

# Load packages
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

@@

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(plant.sp != "Huperzia selago") %>% # we can do without the fern
  left_join(periods) %>% # add periods
  fill(period, .direction = "downup") %>% # fill in periods
  filter(flower.cover > 0) %>% # only count plants in bloom
  dplyr::select(site, date, period, plant.sp, flower.cover) %>%
  mutate(year = factor(year(date))) %>% # convert year to factor
  left_join(fl_traits) # add floral traits

View(filter(survey, is.na(k.type.s)) %>% select(plant.sp) %>% unique())
```

@@

# How is the surveyed pool biased -- phylogenetically, taxonomically, morphologically -- by BB foraging. that is, which taxa/clades/morphologies are over- or underrepresented in the visitation data? And how does this vary with elevation?
```{r}
# By plant family
families.survey <- survey %>% 
  group_by(plant.family) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  mutate(survey.prop = flower.cover/sum(flower.cover))

families.visits <- net %>% 
  group_by(plant.family) %>%
  summarize(visits = n()) %>%
  mutate(visit.prop = visits/sum(visits))

families <- full_join(families.survey, families.visits) %>%
  select(plant.family, survey.prop, visit.prop) %>%
  replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
  #mutate(representation = (visit.prop - survey.prop)/survey.prop) %>% # representation as fold-change
  mutate(representation = visit.prop - survey.prop) %>% # representation as difference
  pivot_longer(cols = c(survey.prop, visit.prop, representation))

ggplot(filter(families, name != "representation"), 
       aes(name, value, fill = plant.family)) +
  geom_col()

ggplot(filter(families, name == "representation"), 
       aes(plant.family, value)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# By Kugler morphotype
morphotypes.survey <- survey %>% 
  group_by(k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  mutate(survey.prop = flower.cover/sum(flower.cover),
         k.type.ss = factor(k.type.ss))

morphotypes.visits <- net %>% 
  group_by(k.type.ss) %>%
  summarize(visits = n()) %>%
  mutate(visit.prop = visits/sum(visits),
         k.type.ss = factor(k.type.ss))

morphotypes <- full_join(morphotypes.survey, morphotypes.visits) %>%
  select(k.type.ss, survey.prop, visit.prop) %>%
  replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
  #mutate(representation = (visit.prop - survey.prop)/survey.prop) %>% # representation as fold-change
  mutate(representation = visit.prop - survey.prop) %>% # representation as difference
  pivot_longer(cols = c(survey.prop, visit.prop, representation))

ggplot(filter(morphotypes, name != "representation"), 
       aes(name, value, fill = k.type.ss)) +
  geom_col()

ggplot(filter(morphotypes, name == "representation"), 
       aes(k.type.ss, value)) +
  geom_col()
```

>>>>

# Load packages
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(mgcv)
library(mgcViz)
library(rstatix)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

@@

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(plant.sp != "Huperzia selago") %>% # we can do without the fern
  left_join(periods) %>% # add periods
  fill(period, .direction = "downup") %>% # fill in periods
  filter(flower.cover > 0) %>% # only count plants in bloom
  dplyr::select(site, date, period, plant.sp, flower.cover) %>%
  mutate(year = factor(year(date))) %>% # convert year to factor
  left_join(fl_traits) # add floral traits
```

@@

# How is the surveyed pool biased -- phylogenetically, taxonomically, morphologically -- by BB foraging. that is, which taxa/clades/morphologies are over- or underrepresented in the visitation data? And how does this vary with elevation?
```{r}
# By plant family
families.survey <- survey %>% 
  group_by(plant.family, site) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  group_by(site) %>%
  mutate(survey.prop = flower.cover/sum(flower.cover))

families.visits <- net %>% 
  group_by(plant.family, site) %>%
  summarize(visits = n()) %>%
  group_by(site) %>%
  mutate(visit.prop = visits/sum(visits))

families <- full_join(families.survey, families.visits) %>%
  select(plant.family, site,  survey.prop, visit.prop) %>%
  replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
  mutate(representation.fold = (visit.prop - survey.prop)/survey.prop, # representation as fold-change
         representation = visit.prop - survey.prop) %>%
  ungroup()

ggplot(families, aes(plant.family, representation)) +
  geom_boxplot()

ggplot(families, aes(plant.family, representation.fold)) +
  geom_boxplot()

aov.fam <- aov(representation ~ plant.family, data = families)
aov.fam.test <- anova_test(aov.fam)
aov.fam.tuk <- tukey_hsd(aov.fam)

aov.fam2 <- aov(representation.fold ~ plant.family, data = filter(families, !is.infinite(representation.fold)))
aov.fam2.test <- anova_test(aov.fam2)
aov.fam2.tuk <- tukey_hsd(aov.fam2)

# By Kugler morphotype
morphotypes.survey <- survey %>% 
  group_by(k.type.ss, site) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  group_by(site) %>%
  mutate(survey.prop = flower.cover/sum(flower.cover),
         k.type.ss = factor(k.type.ss))

morphotypes.visits <- net %>% 
  group_by(k.type.ss, site) %>%
  summarize(visits = n()) %>%
  group_by(site) %>%
  mutate(visit.prop = visits/sum(visits),
         k.type.ss = factor(k.type.ss))

morphotypes <- full_join(morphotypes.survey, morphotypes.visits) %>%
  select(k.type.ss, site,  survey.prop, visit.prop) %>%
  replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
  mutate(representation.fold = (visit.prop - survey.prop)/survey.prop, # representation as fold-change
         representation = visit.prop - survey.prop) %>%
  ungroup()

ggplot(morphotypes, aes(k.type.ss, representation)) +
  geom_boxplot()

ggplot(morphotypes, aes(k.type.ss, representation.fold)) +
  geom_boxplot()

aov.morph <- aov(representation ~ k.type.ss, data = morphotypes)
aov.morph.test <- anova_test(aov.morph)
aov.morph.tuk <- tukey_hsd(aov.morph)

aov.morph2 <- aov(representation.fold ~ k.type.ss, data = filter(morphotypes, !is.infinite(representation.fold)))
aov.morph2.test <- anova_test(aov.morph2)
aov.morph2.tuk <- tukey_hsd(aov.morph2)

library(brms)
brm.morph <- brm(representation ~ k.type.ss, data = morphotypes)
```
